---
title: "Exercises lecture 4 -- joining reshaping tidying"
author: "Paolo Crosetto"
date: "octobre 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)

## import data
df <- flights
```


# `join_...()` family of functions

> first run this setup R code chunk. It will load in your workspace 3 data frames:

- **airports**: avec données sur les aéroports américains
- **flights**: qu'on connait déjà
- **planes**: avec les données pour chaque avion

```{r setup_join}
planes <- planes
flights <- flights
airports <- airports
```

## Exercice 1

> est-ce que les routes plus longues sont desservies par les avions les plus modernes?

*notes*: utilisez `left_join()` et mergez les dataframes `flights` et `planes`

```{r ex11}


```

## Exercice 2

> combien de vols qui partent des trois aéroport de NY atterrisent dans des destinations au dessus de 1000m s.n.m.?

```{r ex12}


```


# tidy data: reshape with `pivot_longer()` and `pivot_wider()`

## Exercice 3

> tidy world_bank_pop dataset so that 'year' is a variable and for each country and each year you have urban population and urban population growh only. Plot as a geom_line the total population for each country over the years.

```{r ex13}

```


